Prediction of Periprosthetic Joint Infection after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden and ...

### 2020-05-09 reports/article.Rmd
m

The cumulative incidence of PJI was estimated using the Kaplan-Meier method. Model development was performed by bootstrap ranking and a logistic least absolute shrinkage and selection operator (LASSO).[@Guo2015; @Baranowski2020] Age and BMI were normalized prior to modeling, and the estimated coefficients thus indicate variable importance on the same scale as categorical variables. 1,000 bootstrap samples were drawn from the observed data set.[@Austin2004] Ten-fold cross validation was performed for every bootstrap sample with a range of potential penalty values ($\lambda$:s) in a logistic LASSO model, keeping only $\lambda$:s that minimized the mean cross-validated deviances in each sample and using these $\lambda$:s to estimate model coefficients for each potential predictor. Potential predictors were identified using piece-wise linear regression and kept as model candidates, regressions were iterated 100 times, and potential predictors that were selected at least once were used in main effects multivariable logistic regression models without penalty and without pre-normalization of numeric variables. Even more parsimonious models with potential predictors chosen at least XXX out of 100 times were evaluated for comparison. Univariable prediction models based on either ASA class, the Charlson or the Elixhauser co-morbidity index were fitted for comparison. Age and BMI were either treated as continuous variables in main effect models, or modelled using restricted cubic splines with two or three knots. In order to account for the potentially U-shaped risk of developing PJI associated with either low or high BMI this variable was also treated as a categorical variable divided into the six BMI intervals suggested by the World Health Organization (WHO; <18.5: underweight, 18.5-24.9: normal, 25-29.9: overweight, 30-34.9: class I obesity, 35.0-39.9: class II obesity, and ???40: class III obesity). Odds ratios for the final models were estimated with 95% confidence intervals.

>>>>

The cumulative incidence of PJI was estimated using the Kaplan-Meier method. Model development was performed by bootstrap ranking and a logistic least absolute shrinkage and selection operator (LASSO).[@Guo2015; @Baranowski2020] Age and BMI were normalized prior to modeling, and the estimated coefficients thus indicate variable importance on the same scale as categorical variables. 1,000 bootstrap samples were drawn from the observed data set.[@Austin2004] Ten-fold cross validation was performed for every bootstrap sample with a range of potential penalty values ($\lambda$:s) in a logistic LASSO model, keeping only $\lambda$:s that minimized the mean cross-validated deviances in each sample and using these $\lambda$:s to estimate model coefficients for each potential predictor. Potential predictors were identified using piece-wise linear regression and kept as model candidates, regressions were iterated 100 times, and potential predictors that were selected at least once were used in main effects multivariable logistic regression models without penalty and without pre-normalization of numeric variables. Even more parsimonious models with potential predictors chosen at least XXX out of 100 times were evaluated for comparison. Univariable prediction models based on either ASA class, the Charlson or the Elixhauser co-morbidity index were fitted for comparison. Age and BMI were either treated as continuous variables in main effect models, or modelled using restricted cubic splines with two or three knots. In order to account for the potentially U-shaped risk of developing PJI associated with either low or high BMI this variable was also treated as a categorical variable divided into the six BMI intervals suggested by the World Health Organization (WHO; <18.5: underweight, 18.5-24.9: normal, 25-29.9: overweight, 30-34.9: class I obesity, 35.0-39.9: class II obesity, and $\geq 40$: class III obesity). Odds ratios for the final models were estimated with 95% confidence intervals.
